### 2017-08-20 sections/conclusion.tex
Follow reviewer's #4 major comments

The goal of this paper is to present the use of SVD decomposition in compression of results from the Finite element method. Implementation details of compression algorithm have been provided, and results of its application on real data have been presented. The implemented algorithm is able to compress arbitrary data using low-rank approximation matrices. If the maximum allowed error was set to $10^{-5}$ then the compression ratio was at most 10\% for all tested results. In many cases compression ratio can be even better -- bellow 1\% of the original size. The important property of compression algorithm is the fact that the approximation error can be set in advance and there is a guarantee that it will not be exceeded.

The main disadvantage is the computational complexity of the compression algorithm. SVD decomposition is very time-consuming operation. However, this operation is performed only once after FEM analysis is finished and before the post-processing is started. Also, randomized version of the SVD decomposition algorithm is much faster and can be used if the slight increase of approximation error is tolerated.

Other complication is the necessity to use special format to store compressed data in the form of SVD decomposition and the post-processor must be updated to use this new format. Post-processor must perform matrix multiplication to get the original results. However, in usual case the data for one analysis step are needed at a time, i.e. only a partial multiplication of a single row is enough. Detailed description of a post-processor implementation is beyond the scope of this paper.

>>>>

The goal of this paper is to present the use of SVD decomposition in compression of results from the Finite element method. Implementation details of the compression algorithm have been provided, and the results of its application on real data have been presented. The implemented algorithm is able to compress arbitrary data using low-rank approximation matrices. If the maximum allowed error was set to $10^{-5}$, then the compression ratio was at most 10\% for all tested results. In many cases compression ratio can be even better -- bellow 1\% of the original size. The important property of the compression algorithm is the fact that the approximation error can be set in advance and there is a guarantee that it will not be exceeded.

The main disadvantage is the computational complexity of the compression algorithm. SVD decomposition is a very time-consuming operation. However, this operation is performed only once after FEM analysis is finished and before the post-processing is started. Also, the randomized version of the SVD decomposition algorithm is much faster and can be used if a slight increase of approximation error is tolerated.

Other complication is the necessity to use a special format to store compressed data in the form of SVD decomposition, and the post-processor must be updated to use this new format. Post-processor must perform matrix multiplication to get the original results. However, in a usual case the data for one analysis step are needed at a time, i.e. only a partial multiplication of a single row is enough. Detailed description of a post-processor implementation is beyond the scope of this paper.

### 2017-08-20 sections/implementation.tex
Update on Overleaf.
Follow reviewer's #4 major comments

Let us assume that the matrix is not empty and is full rank. Then from the formula (\ref{eq:cr-def}) follows that if $r$ equals to the rank of matrix $\mtrx{A}$ the compression ratio is always higher than one. In other words the memory consumption of stored decomposition is bigger than the size of the original matrix. To make the compression algorithm applicable the parameter $r$ must satisfy the condition

@@

\noindent
Considering the usual shape of matrix containing FEM results this inequality is easily satisfiable even for the $r$ being close to the rank of the original matrix as in the typical case the number of nodes or integration points is much higher than the number of analysis steps and therefore $m \ll n$.

\subsection{Algorithm description}
Once the SVD decomposition is calculated the compression algorithm removes a certain number of singular values and corresponding singular vectors. The remaining singular values and vectors represent the compressed data. There are two strategies that influence the way how to get the number of singular values to be preserved -- resulting size and quality. Each strategy is assigned a control parameter that determines compression ratio or approximation error.

@@

\paragraph{Approximation error}
In the usual case the most important measure to take into account is the approximation error. Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined approximation error threshold is not exceeded. To quantify the error the Normalized root-mean-square deviation ($\mathit{NRMSD}$) is used. The normalized error metric enables to work with various data sets that have different scales. $\mathit{NRMSD}$ is defined in section \ref{sec:error}.

To effectively calculate the final rank of the approximation matrix from the desired approximation error the interesting property of singular values

@@

\noindent
where $k=min(m, n)$, i.e. the smallest of two dimensions of the matrix $\mtrx{A}$, is made use of. The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $s_{i}$ of the same matrix $\mtrx{A}$.

@@

Computational complexity of the exact SVD decomposition algorithm is $O(m^2n)$, where $m<n$. This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of execution time on varying matrix dimension is shown. The results of the benchmarks are depicted in Figure \ref{fig:ExeTime_rows} and Figure \ref{fig:ExeTime_columns}. Several observations were made from the results:

\begin{itemize}
\item The algorithm is most efficient in cases where one dimension of the input matrix is very small compared to the other. However, this is almost always the case when compressing results from FEM -- number of incremental or time steps seldom exceeds hundreds.
\item Moreover, incremental or time steps can be split into smaller ranges and the algorithm can be applied on each range separately. This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
\item Randomized SVD algorithm has the same order of algorithmic complexity when full decomposition is required, but yet can significantly reduce execution time. However, the benchmarks are not designed to highlight the benefits of randomized SVD algorithms. The main advantage of the randomized SVD is in the ability to choose the rank of the approximation matrix in advance. In that case only limited number of singular values and corresponding singular vectors is calculated and algorithm performs much faster.
\end{itemize}

Storage size of SVD decomposition itself can also be optimized. $\mtrx{S}$ being a diagonal matrix can be stored as single list of singular values $s_{i}$ or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.

>>>>

Let us assume that the matrix is not empty and is full rank. Then from the formula (\ref{eq:cr-def}) follows that if $r$ equals to the rank of matrix $\mtrx{A}$, the compression ratio is always higher than one. In other words the memory consumption of stored decomposition is bigger than the size of the original matrix. To make the compression algorithm applicable, the parameter $r$ must satisfy the condition

@@

\noindent
Considering the usual shape of matrix containing FEM results, this inequality is easily satisfiable even for the $r$ being close to the rank of the original matrix as in the typical case the number of nodes or integration points is much higher than the number of analysis steps and therefore $m \ll n$.

\subsection{Algorithm description}
Once the SVD decomposition is calculated, the compression algorithm removes a certain number of singular values and corresponding singular vectors. The remaining singular values and vectors represent the compressed data. There are two strategies that influence the way how to get the number of singular values to be preserved -- resulting size and quality. Each strategy is assigned a control parameter that determines compression ratio or approximation error.

@@

\paragraph{Approximation error}
In a usual case, the most important measure to take into account is the approximation error. Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined approximation error threshold is not exceeded. To quantify the error, the Normalized root-mean-square deviation ($\mathit{NRMSD}$) is used. The normalized error metric enables to work with various data sets that have different scales. $\mathit{NRMSD}$ is defined in section \ref{sec:error}.

To effectively calculate the final rank of the approximation matrix from the desired approximation error, the interesting property of singular values

@@

\noindent
where $k=min(m, n)$, i.e. the smallest of two dimensions of the matrix $\mtrx{A}$, is made use of. The above formula states that the sum of squared elements of the matrix $\mtrx{A}$ equals to the sum of squared singular values $s_{i}$ of the same matrix $\mtrx{A}$.

@@

Computational complexity of the exact SVD decomposition algorithm is $\mathrm{O}(m^2n)$, where $m<n$. This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of the execution time on the varying matrix dimension is shown. The results of the benchmarks are depicted in Figure \ref{fig:ExeTime_rows} and Figure \ref{fig:ExeTime_columns}. Several observations were made from the results:

\begin{itemize}
\item The algorithm is most efficient in cases where one dimension of the input matrix is very small compared to the other. However, this is almost always the case when compressing results from FEM -- number of incremental or time steps seldom exceeds hundreds.
\item Moreover, incremental or time steps can be split into smaller ranges and the algorithm can be applied on each range separately. This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
\item The randomized SVD algorithm has the same order of algorithmic complexity when full decomposition is required, but yet can significantly reduce execution time. However, the benchmarks are not designed to highlight the benefits of randomized SVD algorithms. The main advantage of the randomized SVD is in the ability to choose the rank of the approximation matrix in advance. In that case only limited number of singular values and corresponding singular vectors is calculated and algorithm performs much faster.
\end{itemize}

Storage size of SVD decomposition itself can also be optimized. $\mtrx{S}$, being a diagonal matrix, can be stored as single list of singular values $s_{i}$, or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.

### 2017-08-20 sections/math.tex
Follow reviewer's #4 major comments

The objective of a compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation. This saves storage capacity and also accelerates data transfer between computers as the analysis itself and the post-processing of results is usually done on different workstations.

Compression method can be lossy or lossless based on quality of data reconstructed from its compressed representation. Lossless methods are able to fully reconstruct original data. Lossy methods, on the other hand, produce only approximations of original data.

SVD is used as part of the compression algorithm. SVD method applied to arbitrary matrix produces decomposition that consist of corresponding singular values and singular vectors. This process is fully reversible (with the assumption that the numerical errors are negligible). The original matrix can be reconstructed by the multiplication of decomposed parts. However, the compression algorithm is based on modification of decomposition to create low-rank approximation matrix. The reconstructed matrix slightly differs from the original matrix and algorithm therefore performs lossy compression.

@@

The quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $s_{r+1} ...  s_{n}$. The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence. In special cases, when $r=n$, or $s_{i}=0$ for all $i > r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless. In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.

@@

\noindent
The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of left and right singular vectors of matrix $\mtrx{A}$, and a nonnegative, diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$. For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{k})$ floating-point operations, substantially less than classical algorithms.

>>>>

The objective of a compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation. This saves storage capacity and also accelerates the data transfer between computers as the analysis itself and the post-processing of results is usually done on different workstations.

A compression method can be lossy or lossless based on quality of data reconstructed from its compressed representation. Lossless methods are able to fully reconstruct original data. Lossy methods, on the other hand, produce only approximations of original data.

SVD is used as a part of the compression algorithm. The SVD method applied to arbitrary matrix produces decomposition that consist of corresponding singular values and singular vectors. This process is fully reversible (with the assumption that the numerical errors are negligible). The original matrix can be reconstructed by the multiplication of decomposed parts. However, the compression algorithm is based on modification of decomposition to create low-rank approximation matrix. The reconstructed matrix slightly differs from the original matrix and algorithm therefore performs lossy compression.

@@

Quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $s_{r+1} ...  s_{n}$. The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence. In special cases, when $r=n$, or $s_{i}=0$ for all $i > r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless. In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.

@@

\noindent
The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of left and right singular vectors of matrix $\mtrx{A}$, and a nonnegative diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$. For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{k})$ floating-point operations, substantially less than classical algorithms.

### 2017-08-20 sections/results.tex
Follow reviewer's #4 major comments

The number of operations in the compression algorithm with respect to the number of stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the number of operations with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}. Especially on Figure \ref{fig:ExeTime_columns} is clearly visible that the randomized SVD algorithm is much faster than the classical one.

@@

% Execution times
Besides the error also the execution speed of compression algorithm was measured. In Figure \ref{fig:temelin:ExeTime} is comparison of execution times for standard versus randomized SVD compression algorithms. Interestingly, execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with decreasing target rank. If the rank is known ahead of time the fact can be taken advantage of.

@@

The memory consumption of compressed results for reactor containment analysis is summarized in Table \ref{tab:mem-consum}. For different values of compression ratio shows memory size in megabytes. Compression ratio $c$ is an input parameter to the compression algorithm specifying amount of singular values ought to be removed from the SVD decomposition as follows from the equation (\ref{eq:rank-from-comp-ratio}). Size factor describes the final outcome of compression when compared to the original size.

>>>>

The number of operations in the compression algorithm with respect to the number of the stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the number of operations with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}. Especially on Figure \ref{fig:ExeTime_columns} is clearly visible that the randomized SVD algorithm is much faster than the classical one.

@@

% Execution times
Besides the error also the execution speed of compression algorithm was measured. In Figure \ref{fig:temelin:ExeTime}, there is a comparison of execution times for standard versus randomized SVD compression algorithms. Interestingly, execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with decreasing target rank. If the rank is known ahead of time the fact can be taken advantage of.

@@

The memory consumption of compressed results for reactor containment analysis is summarized in Table \ref{tab:mem-consum}. For different values of compression ratio shows memory size in megabytes. In this benchmark the compression ratio $c$ is an input parameter to the compression algorithm. As follows from the equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD decomposition. Size factor describes the final outcome of compression when compared to the original size.
